---
output:
  html_document: default
  pdf_document: default
---
Reproducible Research (course project 1) Markdown File
_______________________________________________________

Loading and preprocessing the data
We assume that the reader set the correct R working directory with the setwd() function.
1.Load the data(i.e read.csv())

```{r}
setwd("~/Desktop")
activity_raw <- read.csv("activity.csv",stringsAsFactors=FALSE)

```
2.Process/transform the data(if necessary)into a format suitable for analysis

```{r}
# Transform the date attribute to an actual date format
activity_raw$date <- as.POSIXct(activity_raw$date, format="%Y-%m-%d")
```

```{r}
#Compute the weekdays from the date attribute
activity_raw <- data.frame(date=activity_raw$date,                                  weekday=tolower(weekdays(activity_raw$date)),
                steps=activity_raw$steps,
                interval=activity_raw$interval)
```

```{r}
## Compute the day type (weekend or weekday)
activity_raw <- cbind(activity_raw,
                      daytype=ifelse(activity_raw$weekday=="saturday"|
               activity_raw$weekday== "sunday", "weekend", "weekday"))
```

```{r}
## Create the final data.frame
activity <- data.frame(date=activity_raw$date,
                       weekday=activity_raw$weekday,
                       daytype=activity_raw$daytype,
                       interval=activity_raw$interval,
                       steps=activity_raw$steps)
```

```{r}
## Clear the workspace
rm(activity_raw)
```
We display the first few wors of the activity data frame:
```{r}
head(activity)
```
What is the mean total number of steps taken per day?
For this part of the assignment, we can ignore the missing values in the dataset.
1.Make a histogram of the total number of steps taken each day

```{r}
# Compute the total number of steps each day (NA values removed)
sum_data <- aggregate(activity$steps, by=list(activity$date),FUN=sum, na.rm=TRUE)
```
# Rename the attributes
names(sum_data) <- c("date", "total")
```
we display the first few rows of the sum_data data frame:
```{r}
head(sum_data)
```
The histogram is given by the following lines of code:
Compute the histogram of the total number of steps each day
  2. Calculate and report the mean and median total number of steps take per day
  The mean and median are computed like
  ```{r}
  mean(sum_data$total)
  median(sum_data$total)
  ```
  These formulas gives a mean and median of 9354 and 10395 respectively.
What is the average daily activity pattern?
  1.Make a time series plot(i.e type="1") of the 5-minute interval(x-axis)and the average number of steps taken,averaged across all days (y-axis)
 
```{r}
 #Clear the workspace
rm(sum_data)
```

```{r}
# compute the means of steps accross all days for each interval
mean_data <- aggregate(activity$steps,
                       by=list(activity$interval),
                       FUN= mean,
                       na.rm= TRUE)
```

```{r}
# Rename the attributes
names(mean_data) <- c("interval", "mean")
```
The time serie plot is created by the following lines of code
 
 ```{r}
 # Compute the time series plot
 plot(mean_data$interval,
      mean_data$mean,
      typle="1",
      col="blue",
      lwd=2,
      xlab="Interval [minutes]",
      ylab="Average number of steps",
      main="Time-series of the average number of steps per intervals/n(NA removed)")
      ```
      
The mean and median are computed like
```{r}
mean(sum_data$total)
median(sum_data$total)
```
These formulas gives a mean and median of 10766 and 10766     respectively.

These values differ greatly from the estimates from the first part of the assignment. The impact of imputing the missing values is to have more data, hence to obtain a bigger mean and median value.

Are there differences in activity patterns between weekdays and weekends?
For this part the weekdays() function may be of some help here.    Use the dataset with the filled-in missing values for this part.
  1.Create a new factor variable in the dataset with two levels - “weekdays” and “weekend” indicating whether a given date is a weekday or weekend day.
# The new factor variable"daytype" was already in the activity data frame
```{r}
head(activity)
```
2.Make a panel plot containing a time series plot (i.e. type = "l") of the 5- minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
# clear the workspace
rm(sum_data)
```

```{r}
# load the lattice graphical library
library(lattice)
```

```{r}
# compute the average number of steps taken, averaged across all daytype avriable
mean_data <- aggregate(activity$steps,
                       by=list(activity$daytype,
                       activity$weekday,activity$interval),mean)
```

```{r}
# Rename the attributes
names(mean_data) <- c("daytype", "weekday","interval", "mean")
```
We display the first few rows of the mean_data data frame:
```{r}
head(mean_data)
```
The time series plot take the following form:

 ```{r}
# compute the time serie plot
 xyplot(mean~interval | daytype, mean_data,
        type="1",
        lwd=1,
        xlab="Interval",
        ylab="Number of steps",
        layout=c(1,2))
```





