---
title: Reproducible Research  Markdown File
author: "Michelle Tan"
date: "8/8/2017"
output: html_document

_______________________________________________________

Loading and preprocessing the data
We assume that the reader set the correct R working directory with the setwd() function.
1.Load the data(i.e read.csv())

```{r}
library(ggplot2)
library(plyr)
setwd("~/Desktop")
activity <- read.csv("activity.csv")
head(activity)
activity$day <- weekdays(as.Date(activity$date))
activity$DateTime <- as.POSIXct(activity$date, format="%Y-%m-%d")
##pulling data without nas
clean <- activity[!is.na(activity$steps),]
head(clean)
```
What is mean total number of steps taken per day?
Calculate the total number of steps taken per day
```{r}
## summarizing total steps per date
sumTable <- aggregate(activity$steps ~ activity$date, FUN= sum,)
colnames(sumTable) <- c("Date", "steps")
```
Make a histogram of the total number of steps taken each day
```{r}
hist(sumTable$steps,breaks=5,xlab="steps", main= "Total steps per day")
```
Calculate and report the mean and median of the total number of steps taken per day
```{r}
## Mean of Steps
as.integer(mean(sumTable$steps))
```
## [1] 10766
```{r}
as.integer(median(sumTable$steps))
```
The average number of steps taken each day was 10766 steps.

The median number of steps taken each day was 10765 steps.
```{r}
clean <- activity[!is.na(activity$steps),]
```
What is the average daily activity pattern?
Make a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
library(ggplot2)
library(plyr)
intervalTable <- ddply(clean, .(interval),summarize ,Avg= mean(steps))
```
Make a plot 
```{r}
P <- ggplot(intervalTable,aes(x=interval,y=Avg),xlab="Interval",ylab="Average Number of Steps")
P +geom_line()+xlab("Interval")+ylab("Average Number of Steps")+ggtitle("Average Number of Steps per Interval")
```
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
##Maximum steps by interval
maxSteps <- max(intervalTable$Avg)
##Which interval contains the maximum average number of steps
intervalTable[intervalTable$Avg==maxSteps,1]
```
The maximum number of steps for a 5-minute interval was 206 steps.

The 5-minute interval which had the maximum number of steps was the 835 interval.

Imputing missing values
Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
##Number of NAs in original data set
nrow(activity[is.na(activity$steps),])
```
The total number of rows with steps = ‘NA’ is 2304.

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

My strategy for filling in NAs will be to substitute the missing steps with the average 5-minute interval based on the day of the week.
```{r}
## Create the average number of steps per weekday and interval
avgTable <- ddply(clean, .(interval, day), summarize, Avg = mean(steps))

## Create dataset with all NAs for substitution
nadata<- activity[is.na(activity$steps),]
## Merge NA data with average weekday interval for substitution
newdata<-merge(nadata, avgTable, by=c("interval", "day"))
```
Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
## Reorder the new substituded data in the same format as clean data set
newdata2<- newdata[,c(6,4,1,2,5)]
colnames(newdata2)<- c("steps", "date", "interval", "day", "DateTime")

##Merge the NA averages and non NA data together
mergeData <- rbind(clean, newdata2)
```
Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
##Create sum of steps per date to compare with step 1
sumTable2 <- aggregate(mergeData$steps ~ mergeData$date, FUN=sum, )
colnames(sumTable2)<- c("Date", "Steps")

## Mean of Steps with NA data taken care of
as.integer(mean(sumTable2$Steps))
```
## [1]10821
```{r}
## Median of Steps with NA data taken care of
as.integer(median(sumTable2$Steps))
```
## [1] 11015
```{r}
hist(sumTable2$Steps, breaks=5, xlab="Steps", main = "Total Steps per Day with NAs Fixed", col="Black")
```



knitr::opts_chunk$set(echo = TRUE)
```
```{r pressure, echo=FALSE}
plot(pressure)
```

